---
title: "Subsetting data with dplyr"
author: "John Little"
subtitle:  "Section 03"
date: "`r Sys.Date()`"
output: html_notebook
---

## Load Library Package

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```


## Import Data

Starwars data set is an on-board dataset with `dplyr`

```{r}
starwars
```

## 5 Verbs

- select
- filter
- arrange
- mutate
- count

    - summarize
    - group_by
    
[slides](https://intro2r.library.duke.edu/slides/)

### Select

1. Select three variables:  name, height, mass
2. Select Helpers:  all variables the end_with("_color")
    
```{r}
starwars %>% 
  select(name, height, mass)

starwars %>% 
  select(ends_with("_color"))     # select helpers
  
starwars %>% 
  select_if(is.numeric)

starwars %>% 
  select(height, weight = mass, films, everything())

starwars %>% 
  select(4, 8:10, 1, everything())
  
```

**YOUR TURN 1 -- 1 minute**

[Exercises from Master the Tidyverse](https://github.com/rstudio/master-the-tidyverse)

- 02-Transform-Data.Rmd

### Filter

```{r}
starwars %>% 
  filter(hair_color == "brown")

starwars %>% 
  filter(mass >= 500)

starwars %>% 
  filter(eye_color %in% c("orange", "blue"))

starwars %>% 
  filter(eye_color == "orange" |
           eye_color == "blue")

starwars %>% 
  filter(eye_color == "brown",   #  &
         hair_color == "brown")

starwars %>% 
  drop_na(mass)

starwars %>% 
  filter(is.na(mass))

starwars %>% 
#  filter(name == "Quarsh Panaka") %>% 
#  filter(str_extract(name, "\\w+") == "Quarsh" ) %>% 
  filter(str_extract(name, "(?<=\\s)\\w+") == "Panaka" )
```

YOUR TURN 2 & 3 -- 5 minutes  (`02-Transform-Data.Rmd`)


## Arrange

```{r}
starwars %>% 
  arrange(height)

starwars %>% 
  arrange(desc(height))

starwars %>% 
  arrange(desc(height), name)

starwars %>% 
  arrange(desc(height), desc(name))
```

YOUR  TURN 4 & 5 -- 4 minutes

## Mutate


```{r}
starwars %>% 
  select_if(is.numeric) %>% 
  mutate(big_number = height + mass + birth_year)
```


```{r message=FALSE, warning=FALSE}
starwars %>% 
  filter(hair_color == "brown") %>% 
  select(name, hair_color, birth_year, height, mass, gender) %>% 
  ggplot(aes(birth_year, mass)) +
  geom_point(aes(color = gender, size = mass)) +
  geom_text(data = . %>% filter(gender == "male"), 
            aes(label = name), position = "dodge", 
            vjust = -1.2, hjust = 0.1) +
  geom_label(data = . %>% filter(gender == "female"), 
            aes(label = name, color = gender), 
            alpha = 0.4,
            position = "dodge", 
            vjust = 1.5, hjust = 0.5) +  
  xlim(0, 230) +
  labs(title = "Star Wars characters with brown hair",
       x = "Birth Year", y = "Mass",
       color = "Gender", alpha = "", size = "Mass",
       caption = "Source:  dplyr::starwars")
```

YOUR TURN -- well NO.  There is no assignment code.

### Summarize

- first let's explore `count`
- then, `group_by`


```{r}
starwars %>% 
  count(hair_color, sort = TRUE)
```

```{r}
starwars %>% 
  group_by(hair_color) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total))

starwars %>% 
  group_by(hair_color, gender) %>% 
  summarise(total = n(), 
            min_height = min(height, na.rm = TRUE), 
            max_height= max(height, na.rm = TRUE)) %>% 
  arrange(desc(total))

starwars %>% 
  group_by(hair_color) %>% 
  summarise(total = n()) %>% 
  summarise(grand_total = sum(total))
```

YOUR TURN -- exercises 6 through 12.


## FOLLOW-UP  

[Working with data](https://rstudio.cloud/learn/primers/2)