---
title: "Tidy Data"
output: html_notebook
---

Adapted from Garrett Grolemund's [Master the tidyverse](https://github.com/rstudio/master-the-tidyverse).  CC BY RStudio.

```{r setup}
library(tidyverse)
library(babynames)

# Toy data
cases <- tribble(
  ~Country, ~"2011", ~"2012", ~"2013",
      "FR",    7000,    6900,    7000,
      "DE",    5800,    6000,    6200,
      "US",   15000,   14000,   13000
)

pollution <- tribble(
       ~city, ~size, ~amount,
  "New York", "large",      23,
  "New York", "small",      14,
    "London", "large",      22,
    "London", "small",      16,
   "Beijing", "large",     121,
   "Beijing", "small",     121
)


data_to_separate <- tribble(
  ~combined_variable,
  "2020_m_15",
  "2020_f_20",
  "2019_m_10",
  "2020_f_15",
  "2018_m_20",
  "2018_f_10"
)



# To avoid a distracting detail during class
names(who) <- stringr::str_replace(names(who), "newrel", "new_rel")
```


## Examples

### Pivot Longer

```{r}
cases

```


```{r}
cases %>% 
  pivot_longer(c("2011", "2012", "2013"),
               names_to = "year",
               values_to = "values")
```


### Pivot Wider

```{r}
pollution 
```


```{r}
pollution %>% 
  pivot_wider(names_from = size,
              values_from = amount)
```

## Separate / Unite

```{r}
data_to_separate
data_to_separate %>% 
  separate(combined_variable, into = c("date", "gender", "value"), sep = "_")
```




## Your Turn 1

On a sheet of paper, draw how the cases data set would look if it had the same values grouped into three columns: **country**, **year**, **n**

## Your Turn 2

Use `pivot_longer` to reorganize `table4a` into three columns: **country**, **year**, and **cases**.

```{r}
table4a %>% 
  pivot_longer(___,
               names_to = "____",
               values_to = "_____")
```

## Your Turn 3

On a sheet of paper, draw how this data set would look if it had the same values grouped into three columns: **city**, **large**, **small**

## Your Turn 4

Use `pivot_wider()` to reorganize `table2` into four columns: **country**, **year**, **cases**, and **population**.

```{r}
table2 %>% 
  pivot_wider(names_from = ____,
              values_from = _____)
```

## Your Turn 5

Pivot longer the 5th through 60th columns of `who` into a key column: value column pair named **codes** and **n**. Then select just the `country`, `year`, `codes` and `value` variables.

```{r}
who %>% 
  pivot_longer(____
               names_to = "____") %>% 
  drop_na(value) %>% 
  select(-iso2, -iso3)


  
```

## Your Turn 6

Separate the `codes` column into **method**, **gender**, and **age** columns.

```{r}

who %>% 
  pivot_longer(5:60,
               names_to = "codes") %>% 
  drop_na(value) %>% 
  select(-iso2, -iso3) %>% 
  separate(codes, into = ________________, sep = " ") %>% 
  select(-new) %>% 
  separate(sexage, into = _______________, sep = 1)
```

## Your Turn 7

Reshape the layout of this data. Calculate the percent of male (or female) children by year. Then plot the percent over time.

```{r}
babynames %>%
  group_by(year, sex) %>% 
  summarise(n = _______) %>% 
  pivot_wider(names_from = _____,
              values_from = ____) %>% 
  mutate(percent_female = F / (M + F) * 100) %>% 
  ggplot() +
  geom_line(aes(x = ___, y = ____))

```

***

# Take Aways

Data comes in many formats but R prefers just one: _tidy data_.

A data set is tidy if and only if:

1. Every variable is in its own column
2. Every observation is in its own row
3. Every value is in its own cell (which follows from the above)

What is a variable and an observation may depend on your immediate goal.

## Follow-up

https://rstudio.cloud/learn/primers/4
