---
title: "Tidy Data"
output: html_notebook
---

Adapted from Garrett Grolemund [Master the tidyverse](https://github.com/rstudio/master-the-tidyverse).  CC BY RStudio

```{r setup}
library(tidyverse)
library(babynames)

# To avoid a distracting detail during class
names(who) <- stringr::str_replace(names(who), "newrel", "new_rel")
```



## Your Turn 1

On a sheet of paper, draw how the cases data set would look if it had the same values grouped into three columns: **country**, **year**, **n**

## Your Turn 2

Use `pivot_longer` to reorganize `table4a` into three columns: **country**, **year**, and **cases**.

```{r}
table4a %>% 
  pivot_longer(2:3,                       #c("1999", "2000"),
               names_to = "year",
               values_to = "cases")
```

## Your Turn 3

On a sheet of paper, draw how this data set would look if it had the same values grouped into three columns: **city**, **large**, **small**

## Your Turn 4

Use `pivot_wider()` to reorganize `table2` into four columns: **country**, **year**, **cases**, and **population**.

```{r}
table2 %>% 
  pivot_wider(names_from = type,
              values_from = count)
```

## Your Turn 5

Pivot longer the 5th through 60th columns of `who` into a key column: value column pair named **codes** and **n**. Then select just the `country`, `year`, `codes` and `value` variables.

```{r}
who %>% 
  pivot_longer(starts_with("new_"),            #5:60,
               names_to = "codes") %>% 
  drop_na(value) %>% 
  select(-iso2, -iso3)


  
```

## Your Turn 6

Separate the `codes` column into **method**, **gender**, and **age** columns.

```{r}

who %>% 
  pivot_longer(5:60,
               names_to = "codes") %>% 
  drop_na(value) %>% 
  select(-iso2, -iso3) %>% 
  separate(codes, into = c("new", "method", "sexage"), sep = "_") %>% 
  select(-new) %>% 
  separate(sexage, into = c("gender","age"), sep = 1)
```

## Your Turn 7

Reshape the layout of this data. Calculate the percent of male (or female) children by year. Then plot the percent over time.

```{r}
babynames %>%
  group_by(year, sex) %>% 
  summarise(n = sum(n)) %>% 
  pivot_wider(names_from = sex,
              values_from = n) %>% 
  mutate(percent_female = F / (M + F) * 100) %>% 
  ggplot() +
  geom_line(aes(year, percent_female))

```

***

# Take Aways

Data comes in many formats but R prefers just one: _tidy data_.

A data set is tidy if and only if:

1. Every variable is in its own column
2. Every observation is in its own row
3. Every value is in its own cell (which follows from the above)

What is a variable and an observation may depend on your immediate goal.

